---
title: "Reproducible Research Assigment 1"
author: "joaoie"
date: "July 18, 2014"
output: html_document
---
#  
# Assignment 1 for Reproducible Research Course  

This is the Assignment 1 for the Reproducible Research assignment 1 written as a R Markdown document.


## First step is to load the data.

1- "activity.csv" is read (from the working directory) into the dataframe "data"  
2- then the date format is corrected.  

```{r}
data<-read.csv("activity.csv")
data$date<-as.Date(data$date,"%Y-%m-%d")
```

## Second step is to determine mean total number of steps taken per day

1- The number of steps per day is determined: the aggregate function aggregates  
the dataframe based on same days, and sums the steps per day.  
The *"interval"* column is then dropped:  

```{r, echo=TRUE}
daysteps<-aggregate(. ~ date, data=data, FUN=sum)
daysteps$interval<-NULL
```

2- Create a Histogram of the total number of steps taken each day:  
 

```{r, echo=TRUE}
ggplot(daysteps, aes(steps)) + geom_histogram(binwidth = 1000, colour="white")
```

3- And calculate the **mean** and **median** total number of steps per day   

```{r, echo=TRUE}
mean<-mean(daysteps$steps)
median<-median(daysteps$steps)
```
The mean is `r mean`, and the median is `r median`.  

## Third step is to determine the average daily activity pattern  

1- The number of steps per 5-minute interval is determined by the aggregate  
function that aggregates the dataframe based on same 5-minute intervals,  
and averages the number of steps per 5-minute interval.   
The *"interval"* column is then dropped:

```{r, echo=TRUE}
x<-data
x$date<-NULL
x<-aggregate(. ~ interval, data=x, FUN=sum)
x$steps<-x$steps/288
plot(x$interval,x$steps, type="l",xlab="5-minute interval of the day",ylab="Average Steps")
```